A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-10-06 latex/main.tex
Note date

@@

>>>>

\subsubsection*{Checking inferred node dates}

To validate the accuracy of times for non-sample internal nodes inserted 
into the ARG
(inferred using tsdate), we compared the dates of first emergence of
Nextstrain clades against the Nextstrain SARS-CoV-2 tree (downloaded on 
2023-01-21). Figure \ref{fig:pango_vs_nextstrain_node_times} shows relatively
close agreement for most clades, with only three (20C, 20E and 21B)
disagreeing by more than 28 days.

Then, to gauge internal consistency of the time estimation procedure, we
selected a subset of sample nodes, matching the distribution of node times 
and average number of children of the non-sample internal nodes. These
selected sample nodes were then re-dated, and the inferred dates compared 
against their known, true sampling dates. The results in Figure 
\ref{fig:node_time_testing} demonstrate that although the inferred and true
times are generally very close, there is a clear bias with re-estimated times
being consistently older (by an average of 37 days). This is in line with 
the findings of other recent
work \cite{sanderson2021chronumental}, and is likely to be a consequence
of edges carrying more mutations than would be expected under a neutral
molecular clock prior, as well as inaccuracies in the recording of 
sample dates.

@@

\begin{figure}
\centering
\includegraphics[width=0.85\linewidth]{node_time_testing.pdf}
\caption{
Heat map showing differences (in days) between actual and re-estimated sample dates for a subset of internal sample nodes. 
}
\label{fig:node_time_testing}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.85\linewidth]{pango_vs_nextstrain_node_times.pdf}
\caption{
Inferred times of internal nodes corresponding to origins of Nextclade clades ($y$-axis) against those estimated in the Nextclade tree ($x$-axis). Confidence intervals shown as horizontal bars. Clades are labelled where dates differ by more than 28 days. 
}
\label{fig:pango_vs_nextstrain_node_times}
\end{figure}
